### 2024-12-14 paper.qmd
update

---
title: "Incorporating spatial distribution at age improves predictive ability of SDMs"
format: pdf
editor: visual
bibliography: references.bib
---

## Abstract

Species distribution models (SDMs) are widely used in fisheries and ecology, providing insights into species' spatial distributions. While many SDMs incorporate external information, such as environmental or habitat features, the majority overlook age-specific information, despite the known influence of age structure on fish distribution and abundance. Including age information may be particularly valuable for species exhibiting age-related spatial patterns driven by ontogenetic shifts, recruitment dynamics, and selective fishing pressures. Here, we develop an age-structured SDM framework using data from the west coast of the USA, and concentrate on two species with distinct life histories: North Pacific hake (*Merluccius productus*) and sablefish (*Anoplopoma fimbria*). Our framework incorporates spatiotemporal models to predict age-specific distributions. We validate our approach using the recursive nature of age data to forecast the spatial distribution of fish by age class across survey locations in subsequent years; these results highlihght that forecast models have predictive skill (more so for sablefish than hake) and the predictive ability is highest for older age classes. Output from our models can be used to predict the spatial age composition, which may be particularly useful for large cohorts (hake from 2010 and 2014 cohorts; sablefish from 2013, 2016, and 2020 cohorts). Our approach is applicable to a wide range of survey types and platforms around the world, and supports the development of spatially explicit management strategies and optimal allocation of fishing effort.

## Introduction

Species distribution models (SDMs) are widely used in marine ecology to understand and predict the spatial distribution of species based on environmental and demographic variables [@melo-merino2020]. These models have informed management decisions for conservation and resource allocation, especially in the face of changing environmental conditions and human impacts, such as fishing pressure [@guisan2005; @elith2009]. Recent advances in SDM applications in the marine environment have incorporated external drivers such as temperature [@fredston2021], habitat [@phillips2017; @laman2018]. A second use of these models has been to predict temporal trends in species biomass from spatiotemporal fisheries independent survey data [@thorson2015]. Trends estimated from SDMs may be included in integrated population models (stock assessments) for commercially fished species, or used to identify risks to populations of conservation concern. Recent developments in these models have included depth [@johnson2019] and habitat variables [@cao2017] to improve trend estimation.

While many SDMs around the world have included variables such as temperature, depth, or habitat, they have generally overlooked age structure, and instead modeled either species occurrences or densities across all age or size classes. Omitting information on age or size may be partially dependent on data availability (ageing fish generally requires sampling otoliths, which increases monetary costs of sampling programs), but may also be driven by the modeling choices of analysts. Much of the previous research around the world has demonstrated that age structure is a fundamental characteristic of fish populations, influencing spatial distribution and abundance due to mechanisms like ontogenetic shifts, recruitment variability, and selective fishing pressures [@rijnsdorp2018]. Ignoring age-specific information in SDMs may lead to incomplete or inaccurate predictions, especially for species with strong age-related spatial gradients.

Incorporating age structure into SDMs may provide more accurate predictions by capturing spatial and temporal variability related to life history traits, age-specific habitat preferences, and population dynamics. Fish often exhibit ontogenetic shifts, where younger individuals occupy different habitats than adults due to factors like predator avoidance, habitat preferences, and diet changes [@werner1984; @ciannelli2004]. These shifts can create age-specific distributions across habitats, particularly for species that undergo distinct life history phases. Additionally, recruitment pulses and migration patterns vary by age, influencing the spatial distribution of younger versus older fish. Fishing selectivity further accentuates these patterns by disproportionately removing older age classes, which alters the population structure and potentially shifts spatial distributions [@hilborn1992; @methot2013]. By incorporating age, SDMs could better predict these dynamics and improve our understanding of species-environment relationships.

Because of the recursive nature of population dynamics, e.g. numbers at age *a* at time *t* can be used to predict numbers at age *a*+1 at time *t*+1, age-structured SDMs with predictive skill also provide a tool for forward prediction, with considerable implications for marine resource management. Reliable predictions of species distributions can help managers anticipate and mitigate issues such as bycatch risk by identifying regions with high densities of specific age cohorts [@lewison2004]. Forecasting the spatial distribution of certain age classes can inform spatially explicit management strategies and help optimize fishing effort to target sustainable cohorts. For example, knowing the expected future locations of specific age classes could aid in creating time-area closures that protect vulnerable life stages or ensure harvests focus on more resilient portions of the population. However, building robust age-structured SDMs requires high-quality, long-term datasets with consistent sampling methods across both age and space to ensure the validity of these predictions.

Around the world, age and life history information is routinely collected by a number of sampling platforms, including from fisheries and fisheries independent surveys. While fisheries dependent data may be used to inform age structured models, sampling from fisheries independent surveys is typically more consistent in space and time. The primary objective of our paper is to explore the utility of developing age - structured SDMs from fisheries independent data. Focusing on two species with contrasting life history characteristics in the North Pacific, the semi-pelagic North Pacific hake (*Merluccius productus*) and longer lived benthic sablefish (*Anoplopoma fimbria*), we first develop a flexible framework for modeling numbers at age. Second, we evaluate the ability of this framework to offer predictive skill in future forecasts, and quantify how the predictive relationship changes as a function of age. Finally, we demonstrate how model output may be used to examine changing species compositions through time, make maps of potential bycatch hotspots, and identify areas with higher concentrations of mature fish as potential targets of fishing effort.

## Methods

### Data

On the west coast of the USA, the West Coast Groundfish Bottom Trawl Survey (WCGBTS) has been an annual survey conduced from 2003 - present. The WCGBTS is designed to estimate the abundance, size, and age composition of groundfish species important to commercial and recreational fisheries found in near-bottom habitats on the west coast of the USA (Keller et al., 2017). The survey effort is concentrated in summer months and has been conducted annually since 2003 (here we use the data through 2018; data are publicly available at https://www.nwfsc.noaa.gov/data). Importantly, the random stratified sampling design, spatial and seasonal coverage, effort, and gears have remained relatively constant within the period we analyse. Like many other surveys around the world, no WCGBTS survey occured in 2020 because of the Covid-19 pandemic. Though the WCGBTS survey samples hundreds of species, we concentrated our analysis on two of the well sampled species from the WCGBTS, North Pacific hake and sablefish. \[more on biological gradients\] While otoliths are sampled continuously during the WCGBTS survey for a broad range of species, otoliths are generally aged when species are being prioritized for stock assessment by the Pacific Fishery Management Council (PFMC); full assessments between species may occur irregularly and be sporadically updated every 5 -- 10 years. Hake and sablefish represent exceptions, as the hake stock is assessed annually by co-managers from the USA and Canada (Grandin et al. 2024) and sablefish is frequently assessed because of its high commercial value. For our analysis, we used hake collected 2007 -- 2019 (mean 650.7 individuals sampled per year) and sablefish collected 2003 -- 2023 (mean 1314.4 individuals sampled per year).

### Spatial age models

For each species, we filtered by sex to focus on females and truncated ages to focus on those with the highest data availability; for hake this included ages 1 -- 5, and for sablefish ages 0 -- 9 (data from these ages represented 54% and 75% of the total aged fish, respectively). For each species - age combination, we constructed a unique spatiotemporal model fit to all years except the last available year for each species. Individual fish were aggregated at the haul level, allowing us to model counts at age. We initially explored modeling counts with a Poisson or Negative Binomial distribution, but as the distributions of numbers at age are dominated by 0s, these models failed to converge. Instead, we summarized the total number of fish sampled and number of fish sampled of the focal age; this allowed counts to be modeled as a binomial family (logit link). Second, we constructed a spatiotemporal model as an extension of Generalized Linear Mixed Models (GLMMs) such that the prediction in location $s$ and time $t$ can be written as

$$
\begin{aligned}
logit \left( p_{\boldsymbol{s},t} \right) &= \boldsymbol{\beta_{t}} + \omega_{\boldsymbol{s}} + \delta_{\boldsymbol{s},t}
\end{aligned}
$$

where $\boldsymbol{\beta_{t}}$ represent time - varying intercepts modeled as a random walk $\boldsymbol{\beta_{t}} \sim N \left( \boldsymbol{\beta_{t-1}}, \sigma_\beta \right)$, the spatial field $\boldsymbol{\omega_{s}} \sim \operatorname{MVNormal} \left( \boldsymbol{0}, \boldsymbol{\Sigma}_\omega \right)$ and the spatiotemporal fields $\delta_{\boldsymbol{s},t}$ are modeled as an AR(1) process $\boldsymbol{\delta}_{t} = \rho \boldsymbol{\delta}_{t-1} + \sqrt{1 - \rho^2} \boldsymbol{\epsilon_{t}}$, where $\boldsymbol{\epsilon_{t}} \sim \operatorname{MVNormal} \left(\boldsymbol{0}, \boldsymbol{\Sigma}_{\epsilon} \right)$

Spatial and spatiotemporal random fields were constructed as Gaussian Markov random fields (GMRFs) using the stochastic partial differential equation approach (SPDE) [@lindgren2011; @lindgren2015]. The SPDE method models the spatial correlation between points as a Matérn covariance function with smoothness parameter $\nu = 1$. Spatial meshes for all ages and species were constructed with a cutoff distance of 50km (this distance controls the spacing of mesh vertices). Parameter estimation was done using the sdmTMB software package [@Anderson2024] with R 4.3.1 [@rcore2024]. The sdmTMB package relies on Template Model Builder (TMB)[@kristensen2016] to quickly and efficiently maximize the marginal log likelihood using auto-differentiation and the Laplace approximation to integrate out random effects. Models were evaluated for convergence (positive-definite Hessian matrix, and a maximum absolute log likelihood gradient \< 0.001) and residuals diagnostics were evaluated with the DHARMa package [@dharma2022].

### Validating future predictive ability

As a first validation, we leveraged the natural recursive element of our data to quantify the ability of our models to predict the future distribution of ages. For each of the species - age models constructed above, we made predictions to the survey locations in the following year (e.g. age 3 hake in years 2007 -- 2018 used to predict the distribution of age 4 hake in 2008 -- 2019). We related the predictions $\hat{p}_{\boldsymbol{s},t}$ to observations by multiplying predictions by the total number of fish sampled for ageing, $\hat{n}_{s,t} = N_{s,t} \cdot \hat{p}_{\boldsymbol{s},t}$. We quantified the relationship between predictions and observations by fitting a simple Poisson GLM, where counts of fish of age $a+1$ in year $t+1$ were treated as the response and $log(\lambda_{s,t}) = \beta_{0} + \beta_{1} \cdot log \left( \hat{n}_{s,t} \right)$. The exponentiated slope parameter $exp(\beta_{1})$ represents a change in expected counts that would be expected from a 1-unit change in $log(\hat{n}_{s,t})$, and can roughly be interpreted as an estimate of relative survival. As a diagnostic check and to understand whether there was a potential time component, we fit a second model as a GLMM with R package glmmTMB; this second model was identical to our Poisson GLM with random effects in $\beta_{0}$ and $\beta_{1}$, where year was used as the grouping variable.

### Using predictions to quantify bycatch risk

To illustrate potential benefits of this kind of modeling, we used the above approach to generate spatiotemporal predictions of age-1 sablefish (from age-0 sablefish the year prior). \[Need some text here on background why sablefish as bycatch is important\]. Gridded predictions from the Poisson GLM represent the expected number of fish per unit of effort in each location. To turn these into a more useful measure of risk, we calculate the total number of age-1 sablefish per unit of effort within a given radius of major fishing ports on the US west coast. For consistency with previous work \[TODO: cite ESR\] we used a radius of 232km.

## Results

One useful output from our spatiotemporal models is the estimated spatial field, $\omega_s$, representing the average spatial distribution across years. Results from our model of North Pacific hake show a concentration of age-1 hake near the coast, and more of latitudinal north - south break for 2 - 3 year old hake (with higher occurrences in the South) @fig-hake-spatial-anomaly. As expected, age 4 hake appear to have a northern distribution, though the gradient is less clear than for ages 2 - 3. For sablefish, our models estimate a concentration of age 0 - 2 sablefish in shallow water, with a deeper distribution by age 3 (@fig-sablefish-spatial-anomaly). Latitudinally, there are several areas in central Oregon that appear to have consistently lower occurrences than average for older age classes (@fig-sablefish-spatial-anomaly).

Estimates of age composition by year may be useful in tracking the distribution of cohorts through time, as well as examining variability in distributions of particular age classes over time. For example, results from our models can be used to project the spatial distributions of strong hake cohorts (2010, 2014, @fig-hake-spatial-composition) and strong sablefish cohorts (2013, 2016, 2020, @fig-sablefish-spatial-composition). These predictions are also useful in revealing subtle differences between strong cohort years. For example, the 2009 cohort of Pacific hake (age 1 in 2010) were predominantly distributed in the northern part of the domain, while the 2013 cohort had hotspots along the entire US west coast. Similarly for sablefish, the 2016 cohort appeared to be more concentrated in the north, versus the more widely distributed 2013 cohort.

When comparing the spatial and spatiotemporal parameters across ages and species, we find that for most ages, there is a decrease in the estimated spatial and spatiotemporal variances as well as a decrease in the spatial range (@fig-spatial-parameters). Spatial and spatiotemporal variances control the magnitude of the peaks and valleys in estimated latent fields, meaning that spatial fields estimated for older fish are less variable. Similarly, the spatial range determines the distance at which locations are functionally independent (for the Matern, $\rho$ \~ 0.13). The decline in range with age implies that spatial distributions also become more concentrated as a function of age (@fig-spatial-parameters).

Our 1-step ahead validation analyses indicate that there is a stronger association between older ages compared to younger ages, and this effect is strongest for sablefish (@fig-glm-coefficients). When comparing our initial Poisson GLM to a more complicated GLMM with random year effects in the intercept and slope, we found little support for shared correlations across age classes of a given species. Shared correlations would indicate for example that age-2 and age-3 sablefish may be more or less predictable in certain years (indicating that information about one age class may be useful in predicting another). With the exception of a strong positive correlation between age-0 and age-1 sablefish (\$\$)

Figure @fig-bycatch-risk aggregates occurrence probabilities for age 0 - 2 sablefish, potentially useful to identify hotspots

Similarly Figure @fig-age5-sablefish shows occurrence probabilities for age 5 sablefish

## Discussion

#### Limitations

-   not predicting abundance (N) -- just proportions
-   We constructed single-age models, and this could be done in a multivariate (multinomial) framework -- but our approach allows for greater flexibility in letting relationships change by age. Tradeoff is that it doesn't impose sum to 1 constraint, binomials are independent

\newpage

```{r label="fig-hake-spatial-anomaly", fig.cap="Estimated spatial anomalies for Pacific hake ages 1 -- 5. These fields represent the average anomalies across all years, 2003 -- 2023.", echo=FALSE}
knitr::include_graphics("plots/hake_spatial_anomalies.png")
```

\newpage

```{r label="fig-sablefish-spatial-anomaly", fig.cap="Estimated spatial anomalies for sablefish ages 0 -- 9. These fields represent the average anomalies across all years, 2003 -- 2023.", echo=FALSE}
knitr::include_graphics("plots/sablefish_spatial_anomalies.png")
```

\newpage

```{r label="fig-hake-spatial-composition", fig.cap="Estimated spatial composition for Pacific hake; rows represent strong cohorts in our dataset (2010, 2014) and columns represent predictions for ages 1 - 5. Full predictions for all cohorts are in the Supplementary Information.", echo=FALSE}
knitr::include_graphics("plots/spatial_composition_subset_Pacific hake.png")
```

\newpage

```{r label="fig-sablefish-spatial-composition", fig.cap="Estimated spatial composition for sablefish; rows represent strong cohorts in our dataset (2013, 2016, 2019) and columns represent predictions for ages 0 - 4. Full predictions for all cohorts are in the Supplementary Information.", echo=FALSE}
knitr::include_graphics("plots/spatial_composition_subset_sablefish.png")
```

\newpage

```{r label="fig-glm-coefficients", fig.cap="Estimated coefficients relating predicted densities of age a fish in year t to observed numbers the following year. The coefficients are interpreted with a multiplicate effect, such that a coefficient of 0.5 translates to an effect of exp(0.5) = 65% increase on the expected count in the future.", echo=FALSE}
knitr::include_graphics("plots/glm_coefficients.png")
```

\clearpage

\newpage

```{r label="fig-bycatch-risk", fig.cap="Estimated probability of occurrence for age 0 -- 2 sablefish.", echo=FALSE}
knitr::include_graphics("plots/sablefish_spatial_risk.png")
```

\newpage

```{r label="fig-age5-sablefish", fig.cap="Estimated probability of occurrence for age 5 sablefish.", echo=FALSE}
knitr::include_graphics("plots/sablefish_spatial_age5.png")
```

\newpage

```{r label="fig-spatial-parameters", fig.cap="Estimated spatial parameters (range, and spatial and spatiotemporal standard deviations). Points represent mean estimates; lines represent 95% confidence intervals", echo=FALSE}
knitr::include_graphics("plots/spatial_parameters.png")
```

\newpage

```{r label="fig-hake-spatial-composition-all", fig.cap="Estimated spatial composition for Pacific hake; rows represent years and columns represent predictions for ages 1 - 5.", echo=FALSE}
knitr::include_graphics("plots/spatial_composition_Pacific hake.png")
```

\newpage

```{r label="fig-sablefish-spatial-composition-all", fig.cap="Estimated spatial composition for sablefish; rows represent years and columns represent predictions for ages 0 - 4.", echo=FALSE}

knitr::include_graphics("plots/spatial_composition_sablefish.png")
```

```{=tex}
\break
\clearpage
```
## References